Edge sign prediction in general graphs and Correlation Clustering

### 2017-09-30 intro/draft.tex
add some refs

The prevalent setting for most of the works discussed above is that the graph connections are: (i)
homogeneous, in the sense that all edges are of a single type, having the same semantics  between
any pair of nodes; and (ii) based on the \emph{homophily} principle~\autocite{Homophily01}, roughly
meaning that nodes are linked because they are similar. However, many real world situations require
a more complex model, with several kind of edges and where nodes can be connected if they are
similar only on some dimensions, or even if they are dissimilar. For instance, an e-commerce website
could let its users express whether they find the reviews written by other users helpful or not. As a
more involved example, consider a social network such as Facebook. Whereas the website offers only a
single \emph{friendship} relation, we may want to distinguish between family ties, work ties and
hobbies ties~\autocite{egoNips12}. Moreover, among one user's \enquote{friends}, there might be some
relations that are actually not positive~\autocite{Yang2012}, whether because of spam, noise or
friendship drift.. Note finally that friendship is only one possible kind of relationship between two
users of Facebook, we could also link them whenever they attend the same event or are tagged in the
same post. Yet another possibility would be to study the same set of users across multiple social
networks~\autocite{mergingNetworks16}, in which case the edge type could denote some common
combinations of connections (\eg Twitter+LinkedIn or Facebook+Instagram). In this thesis, we thus
defend the following statement: \textbf{There exists efficient and accurate methods to predict edge
type in complex networks, reyling either only on the graph topology or also on node metadata.}

>>>>

The prevalent setting for most of the works discussed above is that the graph connections are: (i)
homogeneous, in the sense that all edges are of a single type, having the same semantics  between
any pair of nodes; and (ii) based on the \emph{homophily} principle~\autocite{Homophily01}, roughly
meaning that nodes are linked because they are similar. However, many real world situations require
a more complex model, with several kind of edges and where nodes can be connected if they are
similar only on some dimensions, or even if they are dissimilar. For instance, an e-commerce website
could let its users express whether they find the reviews written by other users helpful or not. As a
more involved example, consider a social network such as Facebook. Whereas the website offers only a
single \emph{friendship} relation, we may want to distinguish between family ties, work ties and
hobbies ties~\autocite{LeskovecEgo12}. Moreover, among one user's \enquote{friends}, there might be some
relations that are actually not positive~\autocite{Yang2012}, whether because of spam, noise or
friendship drift.. Note finally that friendship is only one possible kind of relationship between two
users of Facebook, we could also link them whenever they attend the same event or are tagged in the
same post. Yet another possibility would be to study the same set of users across multiple social
networks~\autocite{mergingNetworks16}, in which case the edge type could denote some common
combinations of connections (\eg Twitter+LinkedIn or Facebook+Instagram). In this thesis, we thus
defend the following statement: \textbf{There exists efficient and accurate methods to predict edge
type in complex networks, reyling either only on the graph topology or also on node metadata.}

### 2017-09-30 troll/related.tex
add some refs

In the first direction, the spectral embedding is illustrated by works from
\textcites{Kunegis2009}{SignedEmbedding15}. We shall describe them more thoroughly in
\autoref{par:cc_spectral}, but here we note that, following the natural orientation of the study of
graph spectrum, they focus more on clustering than \esp{}. Furthermore, the
use of the adjacency matrix usually requires a quadratic running time in the number of nodes, which
makes those methods hardly scalable to large graphs. More recently, there has been great interest in
adapting word embedding techniques such as \textsf{word2vec}~\autocite{word2vec13} to a
\enquote{corpus} of random walks that are considered as documents, while nodes play the role of
word. This allows the unsupervised extraction of node features that can then used to train
downstream classifier, see \autocite{nodeEmbeddingsurvey16} for a recent survey. It can also be
tailored to exploit the specificity of signed graphs.
The goal is to find for every node $u$ a vector $x_u\in\Rbb^d$ in such a way that in this new space,
nodes are close to their positive neighbors and far from their negative neighbors. For instance in
SiNE~\autocite{neuralSigned17}, for a node triplet $u,v,w$ such that the edge $(u,v)$ is positive
and the edge $(u,w)$ is negative, the objective is to find vectors maximizing $f(x_u, x_w) + \delta
- f(x_u, x_v)$. The similarity function $f$ is chosen to be a Siamese multilayers neural network,
whose parameters are learned by back propagation. In SIGNet~\autocite{SIGNet17}, the similarity
function is a sigmoid whose argument $x_u^Tx_v$ is weighted by the sign of the corresponding edge,
with an optimization technique closer to \textsf{word2vec}. Finally, SNE~\autocite{SNE17} uses a
log-bilinear model, and the training objective is to predict, given a path $\mathcal{P}$, which node
should follow $\mathcal{P}$, weighting differently the vector nodes in $\mathcal{P}$ depending of
whether they are the source of a positive or negative edge in $\mathcal{P}$.

>>>>

In the first direction, the spectral embedding is illustrated by works from
\textcites{Kunegis2009}{SignedEmbedding15}. We shall describe them more thoroughly in
\autoref{par:cc_spectral}, but here we note that, following the natural orientation of the study of
graph spectrum, they focus more on clustering than \esp{}. Furthermore, the
use of the adjacency matrix usually requires a quadratic running time in the number of nodes, which
makes those methods hardly scalable to large graphs. More recently, there has been great interest in
adapting word embedding techniques such as \textsf{word2vec}~\autocite{word2vec13} to a
\enquote{corpus} of random walks that are considered as documents, while nodes play the role of
word. This allows the unsupervised extraction of node features that can then used to train
downstream classifier, see
\autocites{nodeEmbeddingsurvey16}{representationLearning17}{graphEmbeddingSurvey17} for three recent
surveys. It can also be
tailored to exploit the specificity of signed graphs.
The goal is to find for every node $u$ a vector $x_u\in\Rbb^d$ in such a way that in this new space,
nodes are close to their positive neighbors and far from their negative neighbors. For instance in
SiNE~\autocite{neuralSigned17}, for a node triplet $u,v,w$ such that the edge $(u,v)$ is positive
and the edge $(u,w)$ is negative, the objective is to find vectors maximizing $f(x_u, x_w) + \delta
- f(x_u, x_v)$. The similarity function $f$ is chosen to be a Siamese multilayers neural network,
whose parameters are learned by back propagation. In SIGNet~\autocite{SIGNet17}, the similarity
function is a sigmoid whose argument $x_u^Tx_v$ is weighted by the sign of the corresponding edge,
with an optimization technique closer to \textsf{word2vec}. Finally, SNE~\autocite{SNE17} uses a
log-bilinear model, and the training objective is to predict, given a path $\mathcal{P}$, which node
should follow $\mathcal{P}$, weighting differently the vector nodes in $\mathcal{P}$ depending of
whether they are the source of a positive or negative edge in $\mathcal{P}$.
